On $K_2$-analogue of Serre problem for $\SL_4$

### 2018-08-18 paper.tex
Writing section of Rehmann theorem

Previously in~\cite{LS17} A.~Lavrenov and the author showed that the local-global principle holds for $\St(4, -)$.
As said before, this together with~\cref{thm:horrocks-sl4} allows us to show the following result, which is an improvement of Tulenbayev stability theorem.
\begin{thm} \label{thm:Serre-problem} %TODO: Fix statement
  For any field $k$ there is an isomorphism $\K_2(4, k[t_1, \ldots, t_n]) \cong \K_2(4, k)$.
  In fact Tulenbayev early stability theorem can be proved starting with the bound $n=4$... 
\end{thm}
It is clear that the positive solution of Serre problem for $\K_2(n, -)$ for $n\geq 4$ is a corollary of this result.
%TODO: Add couple of further corollaries

@@

\section{Preliminaries}

Our notation and conventions follows~\cite[\S~4]{Vav09}.
Let $\Phi$ be an irreducible root system with some fixed basis of simple roots $\Pi = \{\alpha_1, \ldots, \alpha_\ell\}$.
For a root $\alpha\in\Phi$ we denote by $m_i(\alpha)$ the $i$-th coefficient in the expansion of $\alpha$ in $\Pi$,
 i.\,e. $\alpha = \sum_{i=1}^n m_i(\alpha) \alpha_i$.

We denote by $\Phi^\vee$ the \emph{dual root system of $\Phi$} consisting of coroots $\alpha^\vee = 2\alpha/(\alpha, \alpha)$, $\alpha\in \Phi$.
As usual, $P(\Phi^\vee)$ denotes the lattice spanned by the \emph{fundamental weights $\varpi_i$},
 that are uniquely determined by relations $\langle\varpi_i, \alpha_j^\vee \rangle = (\varpi_i, \alpha_j) = \delta_{ij}.$ %TODO: Mistake?

@@

\section{Rehmann--Soul{\'e} theorem}
The aim of this section is to prove a variant of Rehmann--Soul{\'e} theorem on presentation of Steinberg groups.

Let $A$ be arbitrary commutative unital ring and $\mathfrak{m}$ be its ideal.
Denote by $B$ the ring $\mathfrak{m}[t^{-1}] + A[t]$ with the obvious $\mathbb{Z}$-grading.
Clearly, the $k$-th homogeneous component $B_k$ of $B$ equals $A\cdot t^k$ for $k \geq 0$ or $B_k = \mathfrak{m} \cdot t^k$ for $k<0$.

In what follows we will be concerned with finding a compact presentation of the Steinberg group $\St(4, B)$. This presentation is analogous to the that~\cite[Lemma~3.3]{Tu83}, which is formulated for $\St(n, B)$, $n\geq 5$.

@@

There are several reasons why we can't simply refer to the original results of U.~Rehmann and C.~Soul{\'e}~\cite{Re75} or~\cite{RS76}.
The first is that we want to prove Rehmann and Soul{\'e} prove that $G_m \to G_{m+1}$ is an isomorphism for $m\geq 2$, while we want to %establish this fact starting from $m = 1$.
%The second is that they prove their results only in the special case $B = A[t]$, moreover, they make assumptions about the ground ring $A$
%(it is assumed that $A=k$ is a field in~\cite{Re75} and that $A=\mathbb{Z}$ in~\cite{RS76}).
% (this can be replaced by the weaker assumption that $A$ is additively generated by its units).

 \begin{lemma} In the group $G_1$ the following two families of relations are equivalent:
\begin{align} \nonumber \tag{$d_1$}
[[x_{\alpha}(a t), x_{\beta}(b t)],  x_{\gamma}(c)] & = [x_{\alpha + \beta}(a b t),  x_{\gamma}(c t)], & a,b,c \in A,\ \alpha + \beta + \gamma \in \Phi\\
\nonumber \tag{$e_1$}
[[x_{\alpha}(at), x_{\beta}(bt)], x_{\gamma}(ct) ]& = 1, & \alpha + \beta \in \Phi,\ \alpha+\beta \perp \gamma \end{align}
 \end{lemma}

>>>>

Previously in~\cite{LS17} A.~Lavrenov and the author showed that the local-global principle holds for $\St(4, -)$.
As said before, this together with~\cref{thm:horrocks-sl4} allows us to show the following result, which is an improvement of Tulenbayev stability theorem.
\begin{thm} \label{thm:Serre-problem} %TODO: Fix statement
  For any field $k$ there is an isomorphism $\K_2(4, k[t_1, \ldots, t_n]) \cong \K_2(4, k)$.
  In fact Tulenbayev early stability theorem can be proved starting with the bound $n=4$... 
\end{thm}
It is clear that the positive solution of Serre problem for $\K_2(n, -)$ for $n\geq 4$ is a corollary of this result.
%TODO: Add corollary: some building in simply-connected
%TODO: if 2 is invertible and E = SL then H_2(SL_4(R)) = H_2(SL_4(R[t]))

@@

\section{Preliminaries}
\begin{comment}
Our notation and conventions follows~\cite[\S~4]{Vav09}.
Let $\Phi$ be an irreducible root system with some fixed basis of simple roots $\Pi = \{\alpha_1, \ldots, \alpha_\ell\}$.
For a root $\alpha\in\Phi$ we denote by $m_i(\alpha)$ the $i$-th coefficient in the expansion of $\alpha$ in $\Pi$,
 i.\,e. $\alpha = \sum_{i=1}^n m_i(\alpha) \alpha_i$.

We denote by $\Phi^\vee$ the \emph{dual root system of $\Phi$} consisting of coroots $\alpha^\vee = 2\alpha/(\alpha, \alpha)$, $\alpha\in \Phi$.
As usual, $P(\Phi^\vee)$ denotes the lattice spanned by the \emph{fundamental weights $\varpi_i$},
 that are uniquely determined by relations $\langle\varpi_i, \alpha_j^\vee \rangle = (\varpi_i, \alpha_j) = \delta_{ij}.$ %TODO: Mistake?
\end{comment}

@@

\[ T_{ij}(at^{-1}) = \sigma_i T_{ij}(a) \sigma_i^{-1} = \sigma_j^{-1} T_{ij}(a) \sigma_j \]
\[ [[x_{ij}(a t^{-1}), x_{jk}(b t^{-1})],  x_{kl}(c)] = [x_{ik}(a b t^{-1}),  x_{kl}(c t^{-1})] \]
\[ \sigma_i [[x_{ij}(a), x_{jk}(b t^{-1})],  x_{kl}(c)] \sigma_i^{-1} =  \sigma_i [x_{ik}(a b),  x_{kl}(c t^{-1})] \sigma_i^{-1} \]
\begin{multline}
[[x_{ij}(at^{-1}), x_{jk}(bt^{-1})], x_{jl}(ct^{-1})] = \sigma_i [[ x_{ij}(a), x_{jk}(bt^{-1})], x_{jl}(ct^{-1})] \sigma_i^{-1} = \\
\sigma_i \sigma_k^{-1} \sigma_l^{-1} [[ x_{ij}(a), x_{jk}(b)], x_{jl}(c)] \sigma_l \sigma_k \sigma_i^{-1} = 1
\end{multline}
\[ [b, c] =1,\ [a^c, b] = c^{-1} a c b c^{-1} a^{-1} c b^{-1} =  c^{-1} a b  a^{-1} b^{-1} c = [a, b]^c \]

We need to prove that $\sigma_i$ commute with $x_{jk}(a)$ if $j\neq i, k \neq i$.

\section{A variant of Rehmann theorem}
Let $A$ be arbitrary commutative unital ring and $\mathfrak{m}$ be its ideal.
Denote by $B$ the ring $\mathfrak{m}[t^{-1}] + A[t]$ with the obvious $\mathbb{Z}$-grading.
Clearly, the $k$-th homogeneous component $B_k$ of $B$ equals $A\cdot t^k$ for $k \geq 0$ or $B_k = \mathfrak{m} \cdot t^k$ for $k<0$.
If $a$ is a homogeneous element of $B$ we denote its homogeneous degree by $|a|$.

In what follows we will be concerned with finding a presentation of the Steinberg group $\St(4, B)$ in terms of those generators $x_{ij}(\xi)$
 for which $\xi$ is homogeneous of degree $\leq 1$ (i.\,e. $|\xi| \leq 1$).
Our presentation should be thought of as an extension of~\cite[Lemma~3.3]{Tu83} to the case $n=4$ (with variables $t$ and $t^{-1}$ swapped).
Alternatively, our result can be thought of as a variant of the classical theorem of Rehmann~\cite{Re75} (cf. also with Rehmann--Soul{\'e} theorem in~\cite{RS76}).

There are several reasons why we can not simply refer to~\cite{Re75} or~\cite{RS76} and need to repeat arguments from these papers.
\begin{itemize}
 \item Rehmann and Soul{\'e} deal with usual polynomial rings rather than $\mathbb{Z}$-graded rings
  (i.\,e. their result corresponds to the special case $\mathfrak{m}=0$ of our result);
 \item their presentation is formulated in terms of generators $x_{ij}(\xi)$ with $|\xi| \leq 2$ while for our purposes we need a presentation in terms
  of the smaller set of generators with $| \xi |\leq 1$;
 \item it is assumed that $A=k$ is a field in~\cite{Re75} and that $A=\mathbb{Z}$ in~\cite{RS76}, while work with
  arbitrary coefficient ring $A$.
\end{itemize}

@@

 %The second is that they prove their results only in the special case $B = A[t]$, moreover, they make assumptions about the ground ring $A$
%().
% (this can be replaced by the weaker assumption that $A$ is additively generated by its units).

 \begin{lemma} In the group $G_1$ the following two families of relations are equivalent:
\begin{align} 
\nonumber \tag{$d_1$}
[[x_{\alpha}(a t), x_{\beta}(b t)],  x_{\gamma}(c)] & = [x_{\alpha + \beta}(a b t),  x_{\gamma}(c t)], & a,b,c \in A,\ \alpha + \beta + \gamma \in \Phi\\
\nonumber \tag{$df_1$}
[[x_{ij}(a t), x_{jk}(b t)],  x_{kl}(c)] & = [x_{ik}(a b t),  x_{kl}(c t)]\\
\nonumber \tag{$e_1$}
[[x_{\alpha}(at), x_{\beta}(bt)], x_{\gamma}(ct) ]& = 1, & \alpha + \beta \in \Phi,\ \alpha+\beta \perp \gamma \\
\nonumber \tag{$ef_1$}
[[x_{ij}(at), x_{jk}(bt)], x_{jl}(ct) ]& = 1
\end{align}
 \end{lemma}
